---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(Hmisc)
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(corrplot)
library(GGally)
library(tidyr)
library(car)
library(olsrr)
library(MASS)
library(brant)
library(ordinal)
library(nnet)
library(reshape2)
```


Load the wvs dataset, convert -1-.- Don´t know -2-.- No answer -4-.- Not asked -5-.- Missing; Not available 
to NA and then count the number of NAs
```{r}
wvs_raw <- read_excel("C:/Users/Ryan/Academics/Stat 218/wvs.xlsx")

wvs_raw[wvs_raw == -1 | wvs_raw == -2 | wvs_raw == -4 | wvs_raw == -5] <- NA
```

Check Spearman's Correlation of predictor variables to response
```{r}

vars <- c("Q207", "Q287", "Q288R")

wvs_vars <- wvs_raw[ , vars]

cor_results <- sapply(vars[-1], function(var) {
  cor(wvs_vars$Q207, wvs_vars[[var]], use = "pairwise.complete.obs", method = "spearman")
})

cor_sorted <- sort(cor_results, decreasing = TRUE)

print(round(cor_sorted, 3))

cor_df <- data.frame(
  Predictors = names(cor_sorted),
  Correlation = as.numeric(cor_sorted)
)

ggplot(cor_df, aes(x = reorder(Predictors, Correlation), y = Correlation)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = round(Correlation, 2)), hjust = -0.001) +
  labs(title = "Spearman Correlation with Q207",
       x = "Predictor Variables")
```


Convert to proper ordinal factor data types:
```{r}

wvs <- wvs_raw[, c("Q207", "Q287", "Q288R")]

sum(is.na(wvs[c("Q207", "Q287", "Q288R")]))

wvs <- na.omit(wvs[c("Q207", "Q287", "Q288R")])

# Q207 - Social media information source: Ordinal variable
wvs$Q207 <- factor(wvs$Q207, 
                   levels = c(5, 4, 3, 2, 1), 
                   ordered = TRUE)


# Q287 - Social class: Factor with ordered levels
wvs$Q287 <- factor(wvs$Q287, 
                   levels = c(5, 4, 3, 2, 1), 
                   ordered = TRUE)

# Q288R - Income level: Factor with ordered levels
wvs$Q288R <- factor(wvs$Q288R, 
                   levels = c(1, 2, 3), 
                    ordered = TRUE)

str(wvs)
```


# Additional EDA for the Distribution of Age and the Bar plot for Frequency of Social Media Use
```{r}
ggplot(wvs, aes(x = Q207)) + 
  geom_bar(fill = "skyblue") + 
  labs(title = "Frequency of Social Media Use", x = "Usage Level", y = "Count")
```


Perform Ordinal Logistic Regression
```{r}
model <- polr(Q207 ~ Q287 + Q288R, data = wvs, method = "logistic")
summary(model)
```


Test if all the variables hold the Proportional Odds Assumption
```{r}
brant_test <- brant(model)
summary(brant_test)
```


```{r}
sf <- function(y) {
  eps <- 1e-6  # small constant to avoid 0 or 1
  p <- function(k) mean(y >= k)
  qlogis_safe <- function(p) qlogis(pmax(eps, pmin(1 - eps, p)))
  
  c('Y>=1' = qlogis_safe(p(1)),
    'Y>=2' = qlogis_safe(p(2)),
    'Y>=3' = qlogis_safe(p(3)),
    'Y>=4' = qlogis_safe(p(4)),
    'Y>=5' = qlogis_safe(p(5)))
}

```


```{r}
s <- with(wvs, summary(as.numeric(Q207) ~ Q287 + Q288R, fun=sf))
s
```


```{r}
# Logistic regression for Q207 >= 2
model1 <- glm(I(as.numeric(Q207) >= 2) ~ Q287 + Q288R, 
              family = "binomial", data = wvs)

# Logistic regression for Q207 >= 3
model2 <- glm(I(as.numeric(Q207) >= 3) ~ Q287 + Q288R, 
              family = "binomial", data = wvs)

# View model summaries
summary(model1)
summary(model2)

```

For predicted probabilities:
```{r}
# Create all combinations of Q287 and Q288R
newdat <- expand.grid(
  Q287 = factor(1:5, levels = 1:5, ordered = TRUE),
  Q288R = factor(1:3, levels = 1:3, ordered = TRUE)
)

# Add predicted probabilities
newdat <- cbind(newdat, predict(model, newdat, type = "probs"))

# View first few rows
View(newdat)
```

```{r}

# Rename the levels to be interpretable
colnames(newdat)[3:7] <- c("Daily", "Weekly", "Monthly", "LessMonthly", "Never")

# Convert wide to long format
longdat <- melt(newdat, id.vars = c("Q287", "Q288R"),
                variable.name = "Level", value.name = "Probability")

# View
View(longdat)

```


```{r}
ggplot(longdat, aes(x = Q287, y = Probability, fill = Level)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ Q288R, labeller = label_both) +
  labs(title = "Predicted Probabilities of Social Media Use",
       x = "Subjective Social Class",
       y = "Probability") +
  theme_minimal()

```


